A Continual Planner that Switches Between Serial and Decision-Theoretic Planning

### 2010-11-18 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Rough outline for the switching planner part

@@

Finally, we make a number of fairly standard assumptions. First, that
action execution and sensing occurs instantaneously, and that only one
action can be applied at a plan-step. Second, it can arise that in
some propositional states an action is applicable, and in others it is
not. Moreover, a belief-state can assign non-zero probability to
states where that applicability holds, and states where it does
not. We differ
from~\citeauthor{younes:littman:04}~\citeyear{younes:littman:04},
because we forbid the execution of actions that are not applicable in
any state of the current belief.  Also,
unlike~\citeauthor{hoffmann:brafman:2006}~\citeyear{hoffmann:brafman:2006},
we do incorporate a PPDDL-like default semantics, supposing that when
an action \stochAction\ is executed at $\state_i$ where
$\bstate(\state_i) > 0$, that it has no effect on states $\state_i$ if
$\poss(\stochAction) \not\subseteq \state_i$ and $\bstate(\state_i) >
0$.

@@

The modelling language of choice for planning in probabilistic
problems is the Probabilistic Planning Domain Definition
Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
all 3 of the International Planning Competitions since 2004. A
variation on PDDL for describing domains with stochastic actions and
uncertain starting configurations, PPDDL is a declarative first-order
(a.k.a., relational) that facilitates factored description of domains
and problem instances. There are straightforward compilations from
problems expressed compactly in PPDDL to propositional representations
amenable to state-of-the-art planning procedures.  Because this
language cannot model domains that feature partial observability, for
our setting we have developed DTPDDL, an extension of PPDDL to partial
observability.

@@

\scriptsize
\begin{nopagebreak}\begin{tabtt}
\bump \= <observation-def> \bump\=::= \=(:obs\=erve <o-symbol> \+\+\+\+\\
                        %%{[}<extension-def>]  \\
                        :parameters (<typed list (variable)>)  \\
                        <o-def body>) \-\-\- \\
  <o-def body> \> ::= [:state <emptyOr (pre-GD)>] \\
  \> \> [:action <atomic action(term)> ] \\
    \> \> [:effect <emptyOr (o-effect)>] \\
  <o-symbol> \> ::= <name> \\
  <atomic action(t)> \> ::= (<action-symbol> t\zom) \\
  <o-effect> \> ::= (and <c-o-effect>\zom) \\
  <o-effect> \> ::= <c-o-effect> \\
  <c-o-effect> \> ::= \req{:conditional-effects} (when <GD>
                        <o-effect>) \\  
  <c-o-effect> \> ::= \req{:universal-effects} (forall (<typed list (var)>) <o-effect>) \\
  <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic
  <prob> <o-effect>) \\
  <c-o-effect> \> ::=  \req{:probabilistic-effects}\req{:universal-unwinding} \\
%% 
\hspace{3ex}(probabilistic (for-each (<typed list (x)>) <s-f-head>  <o-effect> )) \\
  <c-o-effect> \> ::= <p-o-effect> \\
  <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
  <p-o-effect> \> ::= <atomic o-formula(term)> \\
  <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
  <p-o-effect> \> ::= \req{:fluents}(<assign-op> <o-f-head> <o-f-exp>) \\
  <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
  <o-f-exp> \> ::= <number>\\
  <o-f-exp> \> ::= (<binary-op>  <o-f-exp> <o-f-exp>)\\
  <o-f-exp> \> ::= (<multi-op>  <o-f-exp> <o-f-exp>\oom)\\
  <o-f-exp> \> ::= (- <o-f-exp>)\\
  <o-f-exp> \> ::= <o-f-head>\\
  <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
  <o-f-head> \> ::= <o-function-symbol>\\
\end{tabtt}\end{nopagebreak}
\noteme{<structure-def> ::= <attach-def>}
\normalsize

@@

For example, $(:\pp{predicates} (\pp{location} ?p - VisualObject))$
says that there is a unary predicate

>>>>

\section{Switching Planner Outline}

Outline the high level approach here:
\begin{itemize}
\item Problem input (factored initial state, DTPDDL observations)
\item Determinisation
\item Problem generation for DT
\item Belief revision
\end{itemize}

\section{Determinisation}

\begin{itemize}
\item Generation of deterministic sensing models
\item Determinisation of the initial state
\item Compiling reward model into (soft-)goal model
\item Cost function in FD (maybe this could mentioned in the outline?)
\end{itemize}

\section{Subtask Generation}
\begin{itemize}
\item Goal generation from CP trace
\item Disconfirm actions
\item State pruning according to CP trace
\end{itemize}

\section{Belief revision}
Not sure if this needs its own section
\begin{itemize}
\item Convert inital state into bayesian network
\item Belief updates with observation nodes
\item Create new network for next planning runs
\end{itemize}

@@

Finally, we make a number of fairly standard assumptions. First, that
action execution and sensing occurs instantaneously, and that only one
action can be applied at a plan-step. Second, it can arise that in
some propositional states an action is applicable, and in others it is
not. Moreover, a belief-state can assign non-zero probability to
states where that applicability holds, and states where it does
not. We differ
from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
because we forbid the execution of actions that are not applicable in
any state of the current belief.  Also,
unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
we do incorporate a PPDDL-like default semantics, supposing that when
an action \stochAction\ is executed at $\state_i$ where
$\bstate(\state_i) > 0$, that it has no effect on states $\state_i$ if
$\poss(\stochAction) \not\subseteq \state_i$ and $\bstate(\state_i) >
0$.

@@

The modelling language of choice for planning in probabilistic
problems is the Probabilistic Planning Domain Definition
Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
all 3 of the International Planning Competitions since 2004. A
variation on PDDL for describing domains with stochastic actions and
uncertain starting configurations, PPDDL is a declarative first-order
(a.k.a., relational) language that facilitates factored description of
domains and problem instances. There are straightforward compilations
from problems expressed compactly in PPDDL to propositional
representations amenable to state-of-the-art planning procedures.
Because this language cannot model domains that feature partial
observability, for our setting we have developed DTPDDL, an extension
of PPDDL to partial observability.

Our mobile robot is able to look for a visual object, for example a
box of cornflakes, at a given place. We can model that deterministic
action using the following schema:

\small{
\begin{tabtt}
(\= :action look-for-object \\
  \> :parameters \=(?r - robot ?v - visual-object\\
  \> \> ?l - location) \\
  \> :precondition (and (= (is-in ?r) ?l) ) \\
  \> :effect (and (assign (reward) -3) ) ) \\
\end{tabtt}
}

\noindent That is, the robot can always look for a visual object at
its current location. The only effect this has on the state, is that
the robot incurs a penalty, $3$, that corresponds to the {\em cost} of
performing a visual search. Instances of this operator are useful,
because their application activates an the following sense schema:

\small{
\begin{tabtt}
(\= :sense vision \\
  \> :parameters \=(?r - robot ?v - visual-object\\
  \> \> ?l - location) \\
  \> :execution (look-for-object ?r ?v ?l) \\
  \> :effect ( and (when (= (is-in ?v) ?l) \\
  \> \bump (probabilistic 0.8 \\
  \> \bump (= (observe-is-in ?v) ?l))) \\
  \> (when (not (= (is-in ?v) ?l)) \\
  \> \bump (probabilistic 0.1 \\
  \> \bump (= (observe-is-in ?v) ?l))))) \\
\end{tabtt}
}

\noindent Here, the {\em execution} clause is a lifted description of
the sense action-precondition -- I.e,
$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{desk}) =
\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{desk})$. No
other conditions guard the activation of this sense, therefore if
$(\pp{look-for-object} \pp{Dora} \pp{cornflakes} \pp{bedroom})$ is
executed, we suppose there is a $0.8$ chance of observing the
cornflakes if they are in the $\pp{bedroom}$, and otherwise a $0.1$ of
observing them.\footnote{We have simplified the schemata for
illustrative purposes. In practice, the probabilities of a particular
observation here should be parametrised by the type of visual object
the robot is searching for, and category of the location.}

\scriptsize
\begin{nopagebreak}\begin{tabtt}
<observation-def> \=::= \=(:sense <observation symbol> \\
                        \> \> :parameters (<typed list (variable)>)  \\
                        \> \> <o-def body>) \\
  <o-symbol> \> ::= <name> \\
  <o-def body> \> ::= [:precondition <GD>] \\
  \> \> [:execution <atomic action(term)> ] \\
    \> \> [:effect <o-effect>] \\
  <atomic action(t)> \> ::= (<action symbol> t\zom) \\
  <o-effect> \> ::= (and <c-o-effect>\zom) \\
  <o-effect> \> ::= <c-o-effect> \\
  <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic <prob> <o-effect>) \\
  <c-o-effect> \> ::= <p-o-effect> \\
  <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
  <p-o-effect> \> ::= <atomic o-formula(term)> \\
  <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
  <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
  <o-f-exp> \> ::= <number>\\
  <o-f-exp> \> ::= (- <o-f-exp>)\\
  <o-f-exp> \> ::= <o-f-head>\\
  <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
  <o-f-head> \> ::= <o-function-symbol>\\
\end{tabtt}\end{nopagebreak}
\noteme{<structure-def> ::= <attach-def>}
\normalsize

@@

@@

For example, $(:\pp{predicates} (\pp{location} ?l - VisualObject))$
says that there is a unary predicate
